Visually pleasing knot projections

### 2020-11-06 inst/jma_knot.tex
tweaks

%----------classification, keywords, date
\subjclass{Primary 99Z99; Secondary 00A00}

@@

\begin{abstract}
In this short article I introduce software which creates two
dimensional knot diagrams optimized for visual appearance.  Different
aspects of a knot's appearance are discussed and a framework for
objectively optimizing the visual appeal of a knot projection is
given.

In this short article I show how to create two dimensional knot
diagrams optimized for visual appearance using the R programming
language.  The package associated with this paper is a suite
of software for the creation of production-quality artwork of knot
diagrams, released under GPL2.
\end{abstract}

@@

A {\em mathematical knot} is a smooth, unoriented embedding of a
circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
knots are said to be equivalent if one can be continuously deformed in
to the other; if so, there is a
homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
takes one embedded circle to the other.  In the art world, knot theory
has applications to the study of dance~\cite{khorami2020}, sculpture,
\cite{bosch2010, widmark2020}, and many other branches of mathematical
art~\cite{hart2008}.

@@

@@

Here, knot paths are required to have Bezier handles that are
symmetrically placed with regard to nodes.  Although radius of
curvature is not necessarily matched at either side of a node, visual
continuity of the strand is ensured.  In the diagram, strand crossing
points are far from nodes, as this ensures visual continuity of
strands at crossing points, especially the understrand.

>>>>

%----------classification, keywords, date
\subjclass{Primary 57K10-XX; 32-XX}

@@

\begin{abstract}
In this short article I discuss the \ae sthetics of knot projections
and introduce software which creates two dimensional knot diagrams
optimized for visual appearance.  Different aspects of knot appearance
are discussed and a framework for objectively optimizing the visual
appeal of a knot projection is given.   I use the software to create 
a wide range of knot diagrams.
\end{abstract}

@@

A {\em mathematical knot} is a smooth, unoriented embedding of a
circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
knots are said to be equivalent if the embedding of one can be
continuously deformed to the other; if so, there is a
homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
takes one embedded circle to the other.  Knot theory

In the art world, knot theory
has applications to the study of dance~\cite{khorami2020}, sculpture,
\cite{bosch2010, widmark2020}, and many other branches of mathematical
art~\cite{hart2008}.

@@

One plausible technique for creating knot projections is to consider a
two-dimensional projection of a knot's embedding
$\mathbb{E}\in\mathbb{R}^3$.  However, this approach often results in
poor visual appearance: cusps or other displeasing effects can occur.

@@

Here, knot paths are required to have Bezier handles that are
symmetrically placed with regard to nodes (this is a user-settable
option in inkscape).  Although radius of curvature is not necessarily
matched at either side of a node, visual continuity of the strand is
ensured.  In the diagram, strand crossing points are far from nodes,
as this ensures visual continuity of strands at crossing points,
especially the understrand.
